Responses to Review Comments

### 2025-07-02 writeUps/responses.tex
finished?? editing

\usepackage{xr}
\externaldocument{paper}

\RequirePackage{amsmath,amsfonts,amssymb}
%\RequirePackage[authoryear]{natbib}
\usepackage{longtable}
%\usepackage[notablist,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
\usepackage[doublespacing]{setspace}% To produce a `double spaced' document if required
%\setlength\parindent{24pt}% To increase paragraph indentation when line spacing is doubled

@@

% \theoremstyle{plain}% Theorem-like structures provided by asthma.sty
% \newtheorem{theorem}{Theorem}[section]
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
% \newtheorem{prop}[theorem]{Proposition}

% \theoremstyle{definition}
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem{example}[theorem]{Example}

% \theoremstyle{remark}
% \newtheorem{remark}{Remark}
% \newtheorem{notation}{Notation}

@@

\maketitle
We are grateful for the reviewers' and the editor's careful reading
and comments. We believe that the the substantial changes we made
chages to the manuscript in response to those comments greatly
improved the paper. Most significantly, we clarified our presentation
and added a second application.
We provide point-by-point responses below.

@@

The article effectively presents a simulation study comparing three methods and provides a clear application example of applying GEEPERs. However, there are notable areas for improvement, including a lack of literature review on previous simulation studies about Principal Stratification and some unclear presentations of results. Specific comments are outlined below.

@@

\item \begin{itquote}
     Page 15: The simulation details about how variable Z is simulated and the number of replications for each condition are not clearly explained. Providing more information on these aspects would contribute to the transparency and replicability of the study.
\end{itquote}
We agree that the previous draft was unclear in the description of the outcome-generating model (now equation \ref{eq:y-sim}). We have added material to section \ref{sec:dataGeneration} and Table \ref{tab:factor} that includes the necessary details. Should anything still be unclear, replication code is available in the anonymized GitHub repository at \url{https://anonymous.4open.science/r/psGee-B498/}.

@@

\end{itquote}
We added rationale, including citations, to the paragraph beginning ``Potential and observed outcomes...'' on page \pageref{potential} and to section \ref{sec:manipulatedFactors}.

@@

\item \begin{itquote}      APA Citations and References: The paper does not adhere to the APA 7th style. Authors are advised to consult the APA Formatting and Style Guide for proper citation and reference formatting throughout the manuscript. Consistency in style is crucial for scholarly publications.
\end{itquote}
Thank you. We have fixed the formatting.
\End{enumerate}

@@

\item \begin{itquote}   Please provide more background on M-estimators especially to the extent it evolves to become relevant to the GEEPERs. 
\end{itquote}
Thanks for this suggestion---in response, we added Section \ref{sec:mest}
\item \begin{itquote}      I am wondering if doubly robust PSW estimator would reduce the remaining bias difference between GEEPERs and PSW, and Mixture and PSW.
\end{itquote}
This is an interesting suggestion---you're right that the \psw estimator did not include an outcome model, unlike \geepers and \pmm. However, we doubt that an outcome model would substantially reduce the bias of \psw in the simulation. The advantage of doubly-robust estimation is that an outcome model can correct for model misspecification in the propensity (or principal) score model. However, the principal score model in our simulation study was well-specified---the bias of the \psw estimator was due to an unobserved confounder, $x_3$. Unfortunately, like \psw, doubly-robust estimators rely on ignorability, i.e., that all relevant confounders have been measured and included. 
\item \begin{itquote}      GEEPERs seems great, but also, surprised by how efficient Mixture can be. Are there any mechanisms that link M-estimators to Mixture modeling? Just wondering.
\end{itquote}
We were also surprised by how well the \pmm performed in the simulation, especially given prior literature detailing its vulnerability \citep[][; cited in the paper]{griffin2008application,feller2016principal}. Its similarity to \geepers in some scenarios is less surprising, since they are both essentially mixture model estimators.

\item \begin{itquote}    Would it be possible to offer some evaluation metric for GEEPERs, which would indicate, under some conditions researchers could possibly opt or seek other estimators?
\end{itquote}
This is a great suggestion. We added the paragraph beginning ``The simulation study suggests\dots'' on page \pageref{evalMetric} to give some guidance along those lines. 
\item \begin{itquote}     Quickly checked Frangakis and Rubin (2002). Do you think the effect of bottom-outers (or similar experimental designs) could be partitioned into associative and dissociative effects with GEEPERs?
\end{itquote}
This is correct, but only in a narrow technical sense. 
Under our notation, Frangakis and Rubin (2002) define dissociative effects as contrasts between $\yti$ and $Y_{Ci}$ for the subjects $i$ with $S_{Ci}=\sti$; associative effects are contrasts for subjects with $S_{Ci}\ne\sti$. A contrast between associative and dissociative effects ``measures the degree to which a causal effect of treatment on outcome occurs together with a causal effect of treatment on the surrogate."  Since our paper deals only with the one-way noncompliance case, we set $S_{Ci}=0$ for all $i$ (Assumption 1). In Section \ref{sec:fh2t}, we define $\st=0$ for non-bottom-outers and $\st=1$ for bottom-outers. Therefore, as you wrote, an effect for non-bottom-outers would be dissociative, while an effect for bottom-outers would be associative. However, this is, at least in part, an artifact of our notation. It's true that anyone assigned to control would, by necessity, request fewer than 11 bottom-out hints (the median in the treatment group)---$S_C=0$---and that treatment assignment affects bottom-out-hint usage less for non-bottom-outers than for bottom-outers. However, many non-bottom-outers assigned to treatment will request \emph{some} bottom-out hints, whereas students assigned to control request none---our choice to dichotomize hint requests masks variation that does not exist in the control group. Furthermore, our analysis would have been identical had we set $S_{Ci}$ equal to any other constant or set it to NA.

\item \begin{itquote}    It would be great if you could share the code for estimation and simulation in the next round of review (if there is). Also make sure you have comments in the code so someone reading the article can match the narrative with the code.
\end{itquote}
Full replication code is available at \url{FILL IN} 
\end{enumerate}

@@

\subsection{Simulations}
\begin{itquote}
    In my view, the simulations should show the following:
\begin{itemize}
   \item When the assumptions of extant methods are met, GEEPERS should do no worse.
   \item When the assumptions of extent methods are not met, GEEPERS should do better.
\end{itemize}
\end{itquote}
Depending on the meaning of ``do no worse,'' \geepers may or may not meet those criteria. 
In the scenarios we tested, \geepers is always less biased and its nominal confidence intervals achieve higher empirical coverage than the competitors we tested.
However, the competing methods often boast smaller standard errors than \geepers. 
This set of results is typical for semi- or non-parametric methods, which tend to be less efficient than their parametric counterparts when the latter's assumptions are met, but are, say, less biased when those assumptions are violated.

Unfortunately, \geepers may not even meet that criterion relative to \psw. Our simulation did not include scenarios in which principal ignorability held, but we expect that in such scenarios---depending on other factors---\psw may outperform \geepers in every respect. 
That said, principal ignorability is untestable and, in our opinion, rather strong. 
As we stated on page \pageref{evalMetric} (in the paragraph beginning ``The simulation study suggests''), researchers will have to decide in each application which assumptions are the most plausible, and choose estimators based on that decision.

@@

Following your fine suggestion, we formatted Table \ref{tab:coverage}, highlighting results in red when one or more assumptions have been violated.
This format makes it clearer that, for instance, \pmm makes distributional assumptions while \geepers does not, and that \geepers is less sensitive to the assumptions that both estimators make, especially when $\alpha=0.5$.

@@

\end{document}}

>>>>

\usepackage{xr}
\externaldocument{writeUps/paper}

\RequirePackage{amsmath,amsfonts,amssymb}
%\RequirePackage[authoryear]{natbib}
\usepackage{longtable}

\usepackage[caption=false]{subfig}% Support for small, `sub' figures and tables
\usepackage[doublespacing]{setspace}

@@

@@

\maketitle
We are grateful for the reviewers' and the editor's careful reading
and comments. We believe that the substantial changes we made
to the manuscript in response to those comments greatly
improved the paper. Most significantly, we clarified our presentation
and added a second application.
We provide point-by-point responses below.

@@

@@

\item \begin{itquote}
     Page 15: The simulation details about how variable Z is simulated and the number of replications for each condition are not clearly explained. Providing more information on these aspects would contribute to the transparency and replicability of the study.
\end{itquote}
We agree that the previous draft was unclear in the description of the outcome-generating model (now equation \ref{eq:y-sim}). We have added material to section \ref{sec:dataGeneration} and Table \ref{tab:factor} that includes the necessary details. Should anything still be unclear, replication code is available at \url{https://osf.io/fu2sv/?view_only=b2baebc2f3054ef796717863e4a9662f}.

@@

\end{itquote}
This was a helpful suggestion. We added rationale, including citations, to section \ref{sec:manipulatedFactors}.

@@

\item \begin{itquote}      APA Citations and References: The paper does not adhere to the APA 7th style. Authors are advised to consult the APA Formatting and Style Guide for proper citation and reference formatting throughout the manuscript. Consistency in style is crucial for scholarly publications.
\end{itquote}
Thank you. We have fixed the formatting.
\end{enumerate}

@@

\item \begin{itquote}   Please provide more background on M-estimators, especially to the extent it evolves to become relevant to the GEEPERs. 
\end{itquote}
Thanks for this suggestion---in response, we added Section \ref{sec:mest}
\item \begin{itquote}      I am wondering if doubly robust PSW estimator would reduce the remaining bias difference between GEEPERs and PSW, and Mixture and PSW.
\end{itquote}
This is an interesting suggestion---you're right that the \psw estimator did not include an outcome model, unlike \geepers and \pmm. However, we doubt that an outcome model would substantially reduce the bias of \psw in the simulation. The advantage of doubly-robust estimation is that an outcome model can correct for model misspecification in the propensity (or principal) score model. However, the principal score model in our simulation study was well-specified---the bias of the \psw estimator was due to an unobserved confounder, $x_3$. Unfortunately, like \psw, doubly-robust estimators rely on ignorability, i.e., that all relevant confounders have been measured and included. 
\item \begin{itquote}      GEEPERs seems great, but also, surprised by how efficient Mixture can be. Are there any mechanisms that link M-estimators to Mixture modeling? Just wondering.
\end{itquote}
We were also surprised by how well the \pmm performed in the simulation, especially given prior literature detailing its vulnerability \citep[][; cited in the paper]{griffin2008application,feller2016principal}. Its similarity to \geepers in some scenarios is less surprising, since they are both essentially mixture model estimators.

\item \begin{itquote}    Would it be possible to offer some evaluation metric for GEEPERs, which would indicate, under some conditions researchers could possibly opt or seek other estimators?
\end{itquote}
This is a great suggestion. We added some suggestions at the end of the paragraph beginning ``In sum," on page \pageref{simsum}.

\item \begin{itquote}     Quickly checked Frangakis and Rubin (2002). Do you think the effect of bottom-outers (or similar experimental designs) could be partitioned into associative and dissociative effects with GEEPERs?
\end{itquote}
This is correct, but only in a narrow technical sense. 
Under our notation, Frangakis and Rubin (2002) define dissociative effects as contrasts between $\yti$ and $Y_{Ci}$ for the subjects $i$ with $S_{Ci}=\sti$; associative effects are contrasts for subjects with $S_{Ci}\ne\sti$. A contrast between associative and dissociative effects ``measures the degree to which a causal effect of treatment on outcome occurs together with a causal effect of treatment on the surrogate."  Since our paper deals only with the one-way noncompliance case, we set $S_{Ci}=0$ for all $i$ (Assumption 1). In Section \ref{sec:fh2t}, we define $\st=0$ for non-bottom-outers and $\st=1$ for bottom-outers. Therefore, as you wrote, an effect for non-bottom-outers would be dissociative, while an effect for bottom-outers would be associative. However, this is, at least in part, an artifact of our notation. It's true that anyone assigned to control would, by necessity, request fewer than 11 bottom-out hints (the median in the treatment group)---$S_C=0$---and that treatment assignment affects bottom-out-hint usage less for non-bottom-outers than for bottom-outers. However, many non-bottom-outers assigned to treatment will request \emph{some} bottom-out hints, whereas students assigned to control request none---our choice to dichotomize hint requests masks variation that does not exist in the control group. Furthermore, our analysis would have been identical had we set $S_{Ci}$ equal to any other constant or set it to NA.

\item \begin{itquote}    It would be great if you could share the code for estimation and simulation in the next round of review (if there is). Also make sure you have comments in the code so someone reading the article can match the narrative with the code.
\end{itquote}
\sloppy
Full replication code is available at \url{https://osf.io/fu2sv/?view_only=b2baebc2f3054ef796717863e4a9662f}. 
\end{enumerate}

@@

\subsection{Simulations}
\begin{itquote}
    In my view, the simulations should show the following:
\begin{itemize}
   \item When the assumptions of extant methods are met, GEEPERS should do no worse.
   \item When the assumptions of extent methods are not met, GEEPERS should do better.
\end{itemize}
\end{itquote}
Depending on the meaning of ``do no worse,'' \geepers may or may not meet those criteria. 
In the scenarios we tested, \geepers is always less biased and its nominal confidence intervals achieve higher empirical coverage than the competitors we tested.
However, the competing methods often boast smaller standard errors than \geepers. 
This pattern is typical for semi- or non-parametric methods, which tend to be less efficient than their parametric counterparts when the latter's assumptions are met, but are, say, less biased when those assumptions are violated.

Unfortunately, \geepers may not even meet that criterion relative to \psw. Our simulation did not include scenarios in which principal ignorability held, but we expect that in such scenarios---depending on other factors---\psw may outperform \geepers in every respect. 
That said, principal ignorability is untestable and, in our opinion, rather strong. 
As we stated on page \pageref{simsum} (in the paragraph beginning ``In sum''), researchers will have to decide in each application which assumptions are the most plausible, and choose estimators based on that decision.

@@

Following your fine suggestion, we formatted Tables \ref{tab:factor} and \ref{tab:coverage}, highlighting results in red when one or more assumptions have been violated.
This format makes it clearer that, for instance, \pmm makes distributional assumptions while \geepers does not, and that \geepers is less sensitive to the assumptions that both estimators make, especially when $\alpha=0.5$.

@@

\end{document}
